---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = FALSE,
  asciicast_theme = "pkgdown"
)
asciicast::init_knitr_engine(
  startup = quote({
    library(iddoPal)
    set.seed(1) }),
  echo = TRUE,
  echo_input = FALSE
)
```

# iddoPal

<!-- badges: start -->
<!-- badges: end -->

Colour palettes for R inspired by the [Infectious Diseases Data Observatory](https://www.iddo.org).

This package was written using the `palettes` package, using the accompanying 
[guide](https://mccarthy-m-g.github.io/palettes/articles/creating-packages.html).
See also Gerry Ryan's [`idpalette`](https://github.com/idem-lab/idpalette).

## Included palettes

```{r echo=FALSE, message=FALSE}
library(iddoPal)
plot(iddoPal::iddo_palettes)
```

## Installation

You can install the development version of `iddoPal` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lu-harr/iddoPal")
```

## Usage

```{r}
library(iddoPal)
```

`iddoPal` comes with `r length(iddo_palettes_discrete)` discrete colour palettes and `r length(iddo_palettes_sequential)` sequential colour palettes, which can be accessed from the following R objects:

- `iddo_palettes_discrete` for discrete palettes
- `iddo_palettes_sequential` for sequential palettes
- `iddo_palettes` for all palettes

To preview the palettes in the console simply print them:

```{asciicast iddo-palettes-discrete, eval=FALSE}
# this doesn't render nicely in .Rmd, but try in your own R session:
iddo_palettes_sequential
iddo_palettes_discrete 
```

To preview the palettes in the Plots pane use `plot()`:

```{r plot-iddo-palettes-sequential}
plot(iddo_palettes_sequential)
```

To cast palettes to a tibble use `as_tibble()`:

```{asciicast as-tibble}
as_tibble(iddo_palettes)
```

Palettes can be subset using `[`, `[[`, and `$`.
  
  - To extract a single palette as a colour vector use `[[` or `$`:

    ```{asciicast extract-vector-1}
    iddo_palettes_sequential[["BlWhRd"]]
    ```
    
    ```{asciicast extract-vector-2}
    iddo_palettes_sequential$greys
    ```
    
  - To get names of palettes use `names()`:

    ```{r}
    names(iddo_palettes_sequential)
    ```

## Demonstration

`iddo_palettes` may be then be used for R Base Graphics or `ggplot2` visualisations:

```{r demo}
library(ggplot2)

ggplot(faithfuld,
       aes(waiting, eruptions)) +
  geom_raster(
    aes(fill = density)
  ) +
  scale_fill_gradientn(
    colours = iddo_palettes$BlWhRd
  )

# This isn't very elegant but I promise Base Graphics is usually the most 
# straightforward way of doing things ! I promise !
pal <- iddo_palettes$iddo_new[3:5]

plot(iris$Sepal.Length, iris$Petal.Length, 
     col = alpha(pal[as.numeric(as.factor(iris$Species))], 0.5),
     pch = 16, xlab="Sepal Length", ylab="Petal Length")
legend("bottomright", 
       fill = pal,
       levels(as.factor(iris$Species)))
```

I've found the `BlGyRd` palette to be very handy for cases where we need a diverging palette with clear extremes and midpoint, but I've added `BlGyRdPlus` for even more variation either side of the middle:

```{r demo_plus}
library(ggplot2)
library(dplyr)
library(cowplot)

p1 <- ggplot(faithfuld %>%
         mutate(density = density - min(density),
                density = density / max(density)),
       aes(waiting, eruptions)) +
  geom_raster(
    aes(fill = density)
  ) +
  scale_fill_gradientn(
    colours = iddo_palettes$BlGyRd
  )

p2 <- ggplot(faithfuld %>%
         mutate(density = density - min(density),
                density = density / max(density)),
       aes(waiting, eruptions)) +
  geom_raster(
    aes(fill = density)
  ) +
  scale_fill_gradientn(
    colours = iddo_palettes$BlGyRdPlus
  )

plot_grid(p1, p2)
```

... let's see how that goes

## License

CC0
